### 2017-01-28 manuscript.rmd
added summary statistics and example trees

	#' Add phylogenetic independent contrasts (PIC) with the tree and 
	#' character data in a treeio::treedata object
	#' 
	#' @param gene_trees_pic A list of phylogenetic trees with pics
	#' as treeio::treedata objects 
	#' @return A tibble with combined data, including phylogenetic independent 
	#' contrasts, for internal nodes of all trees
	#' @export
	summarize_contrasts = function( gene_trees_pic ) {
		# Combine the @data slots across all trees. These have labels, node annotations,
		# expression data, contrast results, etc...
		# Also take the absolute value of the contrast, since we only consider magnitude.
		all_nodes = 
			lapply( 
				gene_trees_pic, 
				function( nhx ){
					tags = nhx@data
					tags$gene = digest( nhx ) # Creates a  hash that is unique to each gene tree
					tags %<>% select( -( B ) ) # B has inconsistent types, remove it
					return( tags )
				}
			) %>%
			bind_rows() %>%
			mutate( pic = abs( pic ) )

		# Make a new human readable column for plots that records speciation/ duplication
		all_nodes$Event = NA
		all_nodes$Event[ all_nodes$D == "Y" ] = "Duplication"
		all_nodes$Event[ all_nodes$D == "N" ] = "Speciation"

		# Create a subset that corresponds to nodes with contrasts. This excludes tips.
		nodes_contrast = 
			all_nodes %>%
			filter( ! is.na( pic ) ) %>%
			filter( ! is.na( D ) )

@@

The speciation nodes in the gene tree were then calibrated to the dates specified in the KMRR code. This was necessary to scale the branch lengths so that they are equivalent across the trees, and in turn to scale the phylogenetic independent contrasts (which take branch length into account) so they are equivalent across trees. We used the `chronos()` function from the R package ape [@Paradis:2004dv] for this scaling, with the `correlated` model. See supplementary Materials for additional sensitivity analyses to time calibration. Some trees could not be calibrated with these hard node constraints, and were discarded.

@@

>>>>

	#' Add phylogenetic independent contrasts (PIC) with the tree and 
	#' character data in a treeio::treedata object
	#' 
	#' @param gene_trees_pic A list of phylogenetic trees with pics
	#' as treeio::treedata objects 
	#' @return A tibble with combined data, including phylogenetic independent 
	#' contrasts, for internal nodes of all trees
	#' @export
	summarize_contrasts = function( gene_trees_pic ) {
		# Combine the @data slots across all trees. These have labels, node annotations,
		# expression data, contrast results, etc...
		# Also take the absolute value of the contrast, since we only consider magnitude.
		nodes_all = 
			lapply( 
				gene_trees_pic, 
				function( nhx ){
					tags = nhx@data
					tags$gene = digest( nhx ) # Creates a  hash that is unique to each gene tree
					tags %<>% select( -( B ) ) # B has inconsistent types, remove it
					return( tags )
				}
			) %>%
			bind_rows() %>%
			mutate( pic = abs( pic ) )

		# Make a new human readable column for plots that records speciation/ duplication
		nodes_all$Event = NA
		nodes_all$Event[ nodes_all$D == "Y" ] = "Duplication"
		nodes_all$Event[ nodes_all$D == "N" ] = "Speciation"

		# Create a subset that corresponds to nodes with contrasts. This excludes tips.
		nodes_contrast = 
			nodes_all %>%
			filter( ! is.na( pic ) ) %>%
			filter( ! is.na( D ) )

@@

The gene tree were then time calibrated. The goal is not necesarily to have precise dates for each node, but to scale branch lengths so that they are equivalent across the trees. This in turn scales the phylogenetic independent contrasts (which take branch length into account) so they can be compared appropriately. We found a problem with the clade names on the ENSEMBL Compara trees, which others have also identified [@Daub:2016gd], that first had to be fixed. Hominini is the name for the clade that includes humans and chimps, while Homininae is the clade that includes humans, chimps, and gorillas. Both clades are labeled as Homininae in the Compara trees, though, which would interfere with calibration of these nodes. To fix the problem, we identified all clades labeled Homininae that have no gorilla sequence and renamed them Hominini. We then calibrated the trees by fixing the speciation nodes to the dates specified in the KMRR code, with the exception of Hominini and Homininae. These we set to 7 million years and 9 million years, drawing on the same TimeTree source [@Hedges:2006jv] that KMRR used. We used the `chronos()` function from the R package ape [@Paradis:2004dv] for this calibration, with the `correlated` model. See Supplementary Materials for additional sensitivity analyses to time calibration. Some trees could not be calibrated with these hard node constraints, and were discarded.

@@

#### Summary statistics

```{r contrast_table, cache=TRUE, echo=FALSE, fig.cap="XX_PIC_VAR", warning=FALSE, message=FALSE}

clades_table = 
	nodes_contrast %>% 
	filter(D=="N") %>% 
	.$label %>% 
	table() %>% 
	as.data.frame()

names( clades_table ) = c( "Clade", "N" )

clades_table %<>%  slice( match( calibration_times$clade, Clade) )

knitr::kable(
  clades_table,
  caption = "Supplementary Table 1. Number of speciation nodes with contrasts for each clade on the tree."
)
```

```{r species_table, cache=TRUE, echo=FALSE, fig.cap="XX_PIC_VAR", warning=FALSE, message=FALSE}

	nodes_all = 
			lapply( 
				gene_trees_pic, 
				function( nhx ){
					tags = nhx@data
					tags$gene = digest( nhx ) # Creates a  hash that is unique to each gene tree
					tags %<>% select( -( B ) ) # B has inconsistent types, remove it
					return( tags )
				}
			) %>%
			bind_rows() %>%
			mutate( pic = abs( pic ) )

	species_table = 
		nodes_all %>% 
		filter( ! is.na( species ) ) %>% 
		.$species %>% 
		table() %>% 
		as.data.frame()

	names( species_table ) = c( "Species", "N" )

	species_table %<>% 
		arrange( Species ) %>%
		#mutate( Species = str_to_title( Species ) ) %>%
		mutate( Species = str_replace( Species, "_", " ") )

	knitr::kable(
	  species_table,
	  caption = "Supplementary Table 2. Number of tips for each species in trees that were used to calculate the independent contrasts."
	)

```{r example_trees, cache=TRUE, echo=FALSE, fig.cap="XX_PIC_VAR", warning=FALSE, message=FALSE}

	ntips = lapply( gene_trees_pic, function(x) length( x@phylo$tip.label ))
	to_show = which( ntips < 20 & ntips >15 )[1:8]

	example_trees = lapply( 
		to_show,
		function( x ){
				nhx = gene_trees_pic[[x]]
				nhx@data %<>% mutate( clade_name=label )
				nhx@data$clade_name[ is.tip.nhx(nhx) | nhx@data$D !="N" ] = NA

				nhx@data$Event = NA
				nhx@data$Event[ nhx@data$D == "Y" ] = "Duplication"
				nhx@data$Event[ nhx@data$D == "N" ] = "Speciation"

				nhx %>% 
					ggtree() + 
					geom_tiplab( aes( label=species ), hjust=-0.1 ) + 
					xlim( NA, 1400 ) + 
					geom_tippoint( aes( size=Tau ) ) +
					geom_nodepoint( aes( color=Event ), shape=15, size = 2 ) +
					geom_tippoint( aes( size=Tau ) ) +
					geom_text( aes( label=round( abs( pic ), 3 ) ), vjust=1.5, hjust=1, size=2.5, col="blue" ) + 
					geom_text( aes( label=clade_name ), vjust=-.5, hjust=1, size=2.5, col="black" ) + 
					theme( legend.position="right" )
			}
		)

	print( example_trees )
```

> Supplementary Figure XX_EXAMPLE_TREES | A set of example trees to demonstrate the data and analyses that underlie the results. The value of Tau (expression specificity) is indicated by the sizes of the circles at the tips of the tree. Whether an internal node is a speciation or duplication is indicated by color. Speciation nodes are labeled by clade name. Branch lengths are scaled to time. The blue number is the independent contrast for each node.
